---
title: "Conway's Game of Life in R"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	fig.height = 4,
	fig.width = 6,
	message = FALSE,
	warning = FALSE,
	echo    = FALSE
)
options(width = 300)
```

#### A newbie-friendly R implementation of Conway's Game of Life:  
https://en.wikipedia.org/wiki/Conway's_Game_of_Life    
  
  
`pheatmap` is pretty intuitive and easy to use, but several tools can be used to create the plots

```{r}
library(pheatmap)
```

#### Define the main function which will permute each iteration:  

```{r}
## s_0 is a symmetrical matrix (initial state)
step_fn <- function(s_0) {
  ## create a duplicate matrix for the permuted state (s_1)
  s_1 <- s_0
  ## don't include border cells because it will look nicer
  ## iterate columns
  for(cl in 2:(ncol(s_0)-1)) {
    ## iterate rows
    for(rw in 2:(nrow(s_0)-1)) {
      ## get the value for each of the eight neighbors
      ## around every point
      tl <- s_0[rw-1, cl-1]; t  <- s_0[rw-1, cl] 
      tr <- s_0[rw-1, cl+1]; l  <- s_0[rw, cl-1] 
      r  <- s_0[rw, cl+1]; bl <- s_0[rw+1, cl-1]
      b  <- s_0[rw+1, cl]; br <- s_0[rw+1, cl+1]
      ## calculate how many neighbords are live
      n <- sum(c(tl, t, tr, l, r, bl, b, br) > 0)
      ## permute based on Conway's three rules
      if (s_0[rw, cl] == 1) {
        if (n %in% c(2, 3)) {
        } else {
          s_1[rw, cl] <- 0
        }
      } else if (s_0[rw, cl] == 0) {
        if (n == 3) {
          s_1[rw, cl] <- 1
        }
      }
    }
  }
  ## return the permuted matrix
  return(s_1)
}

```

#### Play the game

```{r}
## set the grid size (256 x 256 in this case)
ngrid <- 256
## number of iterations
steps <- 2000
## create an empty list to store the original state and the permuted states
tbls <- vector(mode = "list", length = steps)
## generate the initial state based on a binomial distribution
## where each cell has a 50% chance of being alive
tbls[[1]] <- matrix(rbinom(ngrid^2, 1, 0.5), ncol = ngrid)
## make the border cells a uniform color (just a cosmetic tweak)
tbls[[1]][1, ] <- rep(0, ngrid); tbls[[1]][, 1] <- rep(0, ngrid)
tbls[[1]][ngrid, ] <- rep(0, ngrid); tbls[[1]][, ngrid] <- rep(0, ngrid)
## iterate the list, applying the permutation function to each item
## and store the permuted matrix at the end of the list
for (step in 1:length(tbls)){ 
  s_0 <- tbls[[step]]
  tbls[[step + 1]] <- step_fn(s_0)
}
```

#### Save the images

```{r}
## iterate the permuted list and call the png() device to save
## the heatmap images to file
for(mat in seq(1, length(tbls))) {
  png(filename = paste0("conway/", mat, ".png"),
      width = 1500, height = 1500, units = "px", 
      pointsize = 12, bg = "white", res = 120)
  pheatmap(tbls[[mat]], cluster_rows = FALSE, cluster_cols = FALSE,
             legend = FALSE, color = c("white", "black"))
  dev.off()
}
```

#### Assemble images into a video  
I used ImageMagick (https://imagemagick.org/index.php), but many tools can be used to do this.  2000 iterations were calculated and compressed to about 13 frames (iterations) per second: 

<iframe width="560" height="315" src="https://www.youtube.com/embed/xKoEEkFOTRA" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

